---
title: "Is there difference in annual income of Male and Female in the USA?"
author: "Radhika Zawar"
output: 
  flexdashboard::flex_dashboard:
    storyboard : TRUE
    vertical_layout: fill
---

<h4 style="font-size:16px;font-family:verdana;">A research data by Bureau of Labor Statistics, January 2015</h4>

<h4 style="font-size:14px;font-family:verdana;" href = "https://www.kaggle.com/jonavery/incomes-by-career-and-gender/data">Data Source: U.S. Incomes by Occupation and Gender. (https://www.kaggle.com/jonavery/incomes-by-career-and-gender/data) </h4>

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyr)
library(stringr)
library(xlsx)
library(plotly)

library(plyr)
library(dplyr)
library(ggplot2)

#loading the datafile
job_role_data <- read.csv("inc_occ_gender.csv", stringsAsFactors = FALSE)

#These three attributes are supposed to be numeric but they are loaded as char so converting them in numeric
job_role_data$All_weekly <- as.numeric(job_role_data$All_weekly)
job_role_data$M_weekly <- as.numeric(job_role_data$M_weekly)
job_role_data$F_weekly <- as.numeric(job_role_data$F_weekly)
sapply(job_role_data,class)

# dropping NA values
job_role_data <- na.omit(job_role_data)

# adding more attributes to simplify the data for the storytelling purpose
#M_yearly is added to provide annual income of Male
job_role_data <- job_role_data %>% mutate(M_yearly = M_weekly*52.1429,
                         F_yearly = F_weekly*52.1429,
                         pay_gap = M_yearly - F_yearly)

# creating new values to be used
Occupation_data <- job_role_data %>%
  select(Occupation,M_weekly,F_weekly, M_yearly,F_yearly,pay_gap) %>%
  filter(str_detect(str_to_upper(Occupation),Occupation))

# ordered by highest annual pay to lowest
Occupation_data_ordered <- Occupation_data[order(Occupation_data$F_yearly),]



```

### Differences in annual income across industries

```{r}

fig <- plot_ly(data = Occupation_data_ordered,
               x = Occupation_data_ordered$M_yearly,
               y = Occupation_data_ordered$Occupation,
               type = 'scatter',
               mode = 'markers',
               name = 'Male',
               color = I("#244CE3"),
               text = Occupation_data_ordered$Occupation,
               textposition = "auto",
               hoverinfo = "text",
               hovertext = paste(
                 "Male Income<br>Industry : ", Occupation_data_ordered$Occupation,
                 "<br> Avg : $", round(Occupation_data_ordered$M_yearly)))


fig <- fig %>% add_trace(data = Occupation_data_ordered,
                         x = Occupation_data_ordered$F_yearly,
                         y = Occupation_data_ordered$Occupation,
                         name = 'Female',
                         color = I("#F445A4"),
                         text = Occupation_data_ordered$Occupation,
                         textposition = "auto",
                         hoverinfo = "text",
                         hovertext = paste(
                           "Female Income<br>Industry : ", Occupation_data_ordered$Occupation,
                           "<br> Avg : $", round(Occupation_data_ordered$F_yearly)))

fig <- fig %>% add_segments(x = Occupation_data_ordered$M_yearly,
                            xend = Occupation_data_ordered$F_yearly,
                            y = Occupation_data_ordered$Occupation,
                            yend = Occupation_data_ordered$Occupation,
                            alpha = 0.5, size = I(1),
                            color = I("black"),
                            text = Occupation_data_ordered$Occupation,
                            textposition = "auto",
                            hoverinfo = "text",
                            hovertext = paste(
                             "Income Gap<br>Industry : ", Occupation_data_ordered$Occupation,
                             "<br> Avg : $", round(Occupation_data_ordered$pay_gap)),
                            showlegend = TRUE,
                            name = "Income Difference")

fig <- fig %>% layout( title = "US Gender Pay difference by Industry, Jan 2015",
                       xaxis = list(title = "Annual pay",
                                   categoryorder = "array",
                                   categoryarray = ~Occupation))

fig <- fig %>% layout(yaxis = list(categoryorder = "array",
                                   categoryarray = Occupation_data_ordered$Occupation))


fig

```


### Average difference in weekly income

```{r}


density1 <- density(job_role_data$F_weekly)

density2 <- density(job_role_data$M_weekly)

gap <- mean(job_role_data$M_weekly) - mean(job_role_data$F_weekly)





fig <- plot_ly(data = job_role_data,
               x = ~density1$x, 
               y = ~density1$y, 
               type = 'scatter', 
               mode = 'none', 
               name = 'Female', 
               fill = 'tozeroy',
               fillcolor = 'rgba(249, 161, 209, 0.5)') 

fig <- fig %>% add_trace(x = ~density2$x, 
                         y = ~density2$y, 
                         name = 'Male', 
                         fill = 'tozeroy',
                         fillcolor = 'rgba(105, 129, 220, 0.5)') 

fig <- fig %>% layout(xaxis = list(title = 'Weekly Income in USD', 
                                   tickvals = c(0,500, 1000, 1500, 2000, 2500), 
                                   ticktext = c("$0","$500", "$1000", "$1500", "$2000", "$2500")),
                      yaxis = list(title = 'Density'),
                      legend = list(x = 1, y = 0.95),
                      title = "General Salary Difference in weekly income of Male and Female (Jan, 2015)")
  
  
fig <- fig %>%  add_segments(x = mean(job_role_data$F_weekly),
                            xend = mean(job_role_data$F_weekly),
                            y = 0,
                            yend = 0.0005,
                            alpha = 1.2, size = I(1),
                            color = I("#F445A4"),
                            name = "Female Average Income",
                            text = mean(Occupation_data_ordered$F_weekly),
                            textposition = "top left",
                            hoverinfo = "text",
                            hovertext = paste(
                             "Female Income<br>Average : $", round(mean(Occupation_data_ordered$F_weekly), digits=2))) 
  
fig <- fig %>%  add_segments(x = mean(job_role_data$M_weekly),
                            xend = mean(job_role_data$M_weekly),
                            y = 0,
                            yend = 0.0005,
                            alpha = 1.2, size = I(1),
                            color = I("#244CE3"),
                            name = "Male Average Income ",
                            text = mean(Occupation_data_ordered$M_weekly),
                            textposition = "top right",
                            hoverinfo = "text",
                            hovertext = paste(
                             "Male Income<br>Average : $", round(mean(Occupation_data_ordered$M_weekly), digits=2))) 


fig

```

### Roles with least difference

```{r}

# ordered by highest annual pay to lowest
Occupation_pay_gap <- job_role_data[order(job_role_data$pay_gap),]

Occupation_pay_gap <- Occupation_pay_gap[1:15,]



fig <- plot_ly(data = Occupation_pay_gap, 
               type = "bar", 
               x=Occupation_pay_gap$pay_gap, 
               y=Occupation_pay_gap$Occupation, orientation="h",
               color = ~pay_gap < 0, colors = c("#244CE3", "#F445A4"), alpha = 0.5,
                name = ~ifelse(pay_gap < 0, "Higher Income for Female", "Higher Income for Male")) 

fig <- fig %>% layout(yaxis = list(categoryorder = "array",
                                   categoryarray = Occupation_pay_gap$Occupation),
                      title = 'Roles where woman have better or almost equal salaries (Jan, 2015)',
                      
                      yaxis = list(title = "Roles"),
                      xaxis = list(title = "Difference in annual income in USD. Left of $0 : Female earn more & Right of $0 : Male earn more",
                      tickvals = c(-5000,-4000, -3000, -2000, -1000, 0, 1000), 
                      ticktext = c("$5000","$4000", "$3000", "$2000", "$1000", "$0", "$1000")),
                      legend = list(x = 0.2, y = 0.95))

fig

```  
 
 <style>

    .storyboard-nav .sbframelist {
        margin: 0 auto;
        width: 94%;
        height: 50px;
        overflow: hidden;
        text-shadow: none;
        margin-bottom: 8px;
    }

    .storyboard-nav .sbnext, .storyboard-nav .sbprev {
        float: left;
        width: 2%;
        height: 50px;
        font-size: 50px;
    }

    </style>